---
title: "README"
author: "Jinju Kim"
date: "6-4-2021"
output: html_document
---

# Computational text analysis for Twitter data 

This package is for text analysis including preprocessing, topic modelling, semantic network, and sentiment analysis. The function includes both analysis and graphic visualization.  

```{r  message=FALSE, warning=FALSE}
library(devtools)
install_github("BeaJJ/ComTxt")

library(ComTxt)
library(dplyr)
library(quanteda)
library(tidyr)
library(scales)
library(igraph)
library(topicmodels)
library(maps)
library(purrr)
library(mallet)
library(stopwords)
library(ggplot2)
```


## Preprocessing 

### 1. geo preprocessing with Premium account

Instead of "exmple_df" place your Data name. Multi_names indicate other names that user might write. I found users wrote Croatia mainly but some people also wrote republic of croatia. You can check from the geo_preprocess function.

```{r message=FALSE, warning=FALSE}
## put all different names of your country check the value of multi_name
?geo_preprocess
example_df <- geo_preprocess(example_df, location_country = "spain" , ##"finland" ##"croatia"
                         multi_name = c("españa","spagna")) ##c("Suomi")  c("republic of croatia")

```

#### Geo information Description 

```{r message=FALSE, warning=FALSE}
geo_map(example_df, location_country = "Spain")
geo_table(example_df)
```

### 3. text preprocessing

Please check the function and ud_lang, stopwords you can choose according to your language. Select your own langauge for preprocessing 

```{r message=FALSE, warning=FALSE}

unique(example_df$lang) 
example_lang.1 <- example_df[example_df$lang == "es", ] 
example_lang.2 <- example_df[example_df$lang == "ca", ]

?twitter_preprocess
## preprocess : lemmatization, stopwords, urls, emoticons
stop_words <- c(stopwords::stopwords("es", source = "stopwords-iso"))
df_lang.1 <- twitter_preprocess(example_lang.1, ud_lang = "spanish", stop_words)

## stopwords 
stop_words <- c(stopwords::stopwords("ca", source = "stopwords-iso"))
df_lang.2 <- twitter_preprocess(example_lang.2, ud_lang = "catalan", stop_words)

df_pre <- rbind(df_lang.1, df_lang.2)
```

## Twitter Description 

### 1. Key words Description 

```{r message=FALSE, warning=FALSE}
keyword_network(df_pre, keyword_remove = "cultura", top_n = 40)
keyword_table(df_pre, keyword_remove= "cultura", top_n = 10)
##select your keyword to zoom in
keyword_network_hub(df_pre,hub="libro", top_n = 40 )
shiny_keyword(df_pre)
```

### 2. Hashtags Description 

```{r message=FALSE, warning=FALSE}
hashtags_network(df_pre, hashtag_remove = "#cultura", top_n = 40)
hashtags_table(df_pre, hashtag_remove = "#cultura", top_n = 10)
##cine #historia #literatura #música
##select your own key hashtags fromm  table
hashtags_network_hub(df_pre, hub = "#arte", top_n = 10)
```

### 3. User Description 

Top n users based on follower, friends, tweets number 

```{r message=FALSE, warning=FALSE}
user_table(df_pre, top_n = 10 )
mention_table(df_pre, top_n = 10)
mention_network(df_pre, top_n = 20)

shiny_mention(df_pre)
```

## Mallet Topic Modeling  

### 1. Topic number K 
n_topic is the topic number K you select. 

```{r message=FALSE, warning=FALSE}
## Topic modeling------------------------------------------------------
?topic_number
topic_number(df_pre)

```

### 2. Topic modelling

```{r}
### mallet topic modeling
?twitter_topic
df_topics <- twitter_topic(df_pre, n_topic = 7)
```

### 3. Topic modelling visualization

```{r}
### data frame words topic
?topic_getwords
topic_getwords(df_topics, n_words = 100)

### draw wordscloud for each topics 3*3 = 9 spots (depending on your topic numbers)
?topic_wordcloud
topic_wordcloud(df_topics, maximum_n = 200, cloud_scale = 2.5)

### draw top n words in each topic 
?topic_wordplot 
topic_wordplot(df_topics,  num_words = 10)

### draw topic radar map based on year distribution 
?topic_radarmap
topic_radarmap(df_pre, df_topics)

?shiny_topic
shiny_topic(df_topics, df_pre)
```

