---
title: 'The space of lifestyles in Finland, 2007 and 2018: Continuity and change'
author: "Ossi Sirkka, Semi Purhonen"
date: '`r format(Sys.Date(), "%d. %B %Y")`'
abstract: "Have the ways in which cultural practices are socially structured changed over the last 10–15 years, a period marked by digitization of cultural consumption and production and increased socio-political polarization? This work-in-progress paper tackles this question by analyzing two repeated nationally representative surveys, collected in Finland in 2007 (N=1,388) and 2018 (N=1,425), allowing comparisons over time and including a wide set of indicators of cultural practices. Applying multiple correspondence analysis (MCA), the paper sets out to construct the space of lifestyles – or, “cultural map” – in Finland in both 2007 and 2018. The analysis carried out, inspired by the few studies having used MCA in analyzing temporal changes, includes three steps. First, the space of lifestyles for 2007 is constructed using more than 100 indicators of cultural practices, while the socio-political divisions are projected onto the space as supplementary variables. Second, the space of lifestyles for 2018 is constructed using exactly identical set of indicators as for 2007 to investigate the nature of changes, if measured symmetrically. Third, to reveal the impact of new cultural practices that were not available for the 2007 data, the space of lifestyles for 2018 is constructed again, this time using also a comprehensive set of additional indicators of social media and other online and offline everyday activities. The results are discussed in the light of other recent research focusing on the temporal dynamics of cultural stratification in different national contexts."
keywords: "Cultural practices, space of lifestyles, taste, stratification, change, multiple correspondence analysis"
output: 
  bookdown::html_document2:
  number_sections: TRUE
  fig_caption: TRUE
  pdf_document:
    toc: TRUE
bibliography: space-of-lifestyles.bib
---

```{r BIBTEXTAGS, include = FALSE}
bib2df::bib2df("space-of-lifestyles.bib")$BIBTEXKEY
```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=14,
                      dev = "png",
                      fig.path=paste0(getwd(),'/images_for_article/'),
                      echo=FALSE, warning=FALSE, message=FALSE)



```

```{r, child = "models_and_data.Rmd"}
```

```{r, child = "introduction_theory.Rmd"}
```

```{r, child = "research_desing.Rmd"}
```

```{r, child = "analysis.Rmd"}
```

# Conclusions

  * Two dimensions are sufficient to represent space of lifestyles in Finland, both 2007 and 2018
    * Both dimensions remained quite stable during 11 year period
    * With identical set of indicators in 2007 and 2018, the explanatory power of traditional-modern axis declined, possible because of 1) cohort effect (most traditional age group got off the survey) and 2) no emerging indicators of contemporary lifestyle practices were included (e.g. gaming, online indicators), which leaded to underpresentation of cultural goods consumed by youngest age cohort, 18-24 year old young adults. This can be confirmed from Figure 3.7, where age group of 25-34 year old adults are as modern as 18-24 year old, which was not the case in 2007. Adding cultural good consumed by youngest population to the analysis, the youngest cohort became also the most modern. 
    * Considering individual items used to construct the space of lifestyles in 2007 and 2018, some cultural indicators experienced notable change in their location in the space. For example, reading at least some books turned from not being separating in neither axis in 2007 to being modern and legitimate practice.
  
  * Conducting MCA with comprehensive set of indicators revealed position of contemporary lifestyle practices in the space of lifestyles. From the indicators available in the survey but not yet included in first two steps of the analysis, majority were contributing more strongly in second, traditional-modern axis. 
  * Being involved in gaming were indicator of modern and slightly illegitimate lifestyle. Activities online and in social media were located also in modern side of the space, but more strongly in the legitimate side of the space.  

# References
# Appendix

## Appendix 1

```{r, appendix1, eval = TRUE}
tables_2007 <- lapply(vanha$data, table, exclude = TRUE)
tables_2018 <- lapply(uusi$data, table, exclude = TRUE)

tables_list <- lapply(seq_len(length(tables_2007)), function(i){
  
  if (all(which(c(length(tables_2007[[i]]), length(tables_2018[[i]])) == min(c(length(tables_2007[[i]]), length(tables_2018[[i]])))) == 1)){
    tables_2007[[i]][length(tables_2007[[i]]) + 1] <- NA
  } else if (all(which(c(length(tables_2007[[i]]), length(tables_2018[[i]])) == min(c(length(tables_2007[[i]]), length(tables_2018[[i]])))) == 2)){
  tables_2018[[i]][length(tables_2018[[i]]) + 1] <- NA
}
    
  
  data.frame(
    Variable = c(str_sub(names(tables_2007[[i]][1]), start = 1, end = nchar(names(tables_2007[[i]][1]))-1 ),
                 rep(" ", length(names(tables_2007[[i]]))-1)),
    Label = names(tables_2007[[i]]),
    Frequency_2007 = c(unname(tables_2007[[i]])),
    Frequency_2018 = c(unname(tables_2018[[i]])))
})

tables_df <- do.call(rbind, tables_list)
tables_df <- tables_df[!str_detect(tables_df$Label, "NA") & !is.na(tables_df$Label),]
tables_df <- add_column(tables_df,
                        Class = c(mod_luokitukset, rep(" ", nrow(tables_df) - length(mod_luokitukset))),
                        .before = "Frequency_2007")


create_kable(tables_df, otsikko = "Full list of indicators used to constuct identical spaces of lifestyles in 2007 and 2018 ")
```

## Appendix 2

```{r, appendix2, eval = TRUE}
new_vars_tables <- lapply(newdat$data[, which(names(newdat$data) == "Toimintapelit"):ncol(newdat$data[,!names(newdat$data) %in% sup.vars])], table, exclude = TRUE)
tables_new_list <- lapply(seq_len(length(new_vars_tables)), function(i){
  data.frame(
    Variable = c(str_sub(names(new_vars_tables[[i]][1]), start = 1, end = nchar(names(new_vars_tables[[i]][1]))-1 ),
                 rep(" ", length(names(new_vars_tables[[i]]))-1)),
    Label = names(new_vars_tables[[i]]),
    Frequency = c(unname(new_vars_tables[[i]])))
})

new_tables_df <- do.call(rbind, tables_new_list)
new_tables_df <- add_column(new_tables_df[str_detect(new_tables_df$Label, "NA", negate = TRUE),], Class = mod_luokitukset_uudet_muuttujat, .after = "Label")


create_kable(new_tables_df, otsikko = "List of additional indicators used to constuct comprehensive space of lifestyles with 2018 data")

```

# References